environmental_data_preperation

### 2018-09-19 src/biological_data_prep.Rmd
amo data extraction

shift the amo_sample column down by 1...
```{r}
library(useful)
amo_yrmt <- shift.column(data = amo_yrmt, columns = "amo_sample", len = 1)
amo_yrmt
```

Note the shifted values are in a new column called amo_sample.Shifted. rename to amo_prev
```{r}
colnames(amo_yrmt)[colnames(amo_yrmt)=="amo_sample.Shifted"] <- "amo_prev"
amo_yrmt
```

```{r}
data_aea <- merge(x = data_aea, y = amoshift[ , c("year", "month", "amo_prev")], by = c("year", "month"), all.x = TRUE)
data_aea
```

@@

You need to extract the seasonal values from the amo dataset
```{r}

```

```{r}
amo_yrmt <- shift.column(data = amo_yrmt, columns = "amo_sample", len = 1)
amo_yrmt
```

>>>>

shift the amo_sample column down by 1...
```{r}
library(useful)
amo_yrmt <- shift.column(data = amo_yrmt, columns = "amo_sample", up = FALSE)
amo_yrmt
```

Note the shifted values are in a new column called amo_sample.Shifted. rename to amo_prev
```{r}
colnames(amo_yrmt)[colnames(amo_yrmt)=="amo_sample.Shifted"] <- "amo_prev"
amo_yrmt
write.csv(amo_yrmt, "../output/env/amo_prev.csv", row.names = FALSE) #for some reason i need to write to a csv and reload for the next step to work properly
```

now grab the data...
```{r}
amo_prev <- read.csv("../output/env/amo_prev.csv")
data_aea <- merge(x = data_aea, y = amo_prev[ , c("year", "month", "amo_prev")], by = c("year", "month"), all.x = TRUE)
data_aea
```

@@

You need to extract the seasonal values from the amo dataset
```{r}
amo_winter <- subset(amo, select = c("year", "WinterAvg"))
amo_winter
```

now shift the winter average data 
```{r}
amo_winter <- shift.column(data = amo_winter, columns = "WinterAvg", up = FALSE)
amo_winter
write.csv(amo_winter, "../output/env/amo_winter.csv", row.names = FALSE) #for some reason need to save to csv then reload in the next step...
```
remember your shifted column is called WinterAvg.Shifted

now extract the WinterAvg.Shifted values to the species dataset
```{r}
amo_winter <- read.csv("../output/env/amo_winter.csv")
data_aea <- merge(x = data_aea, y = amo_winter[ , c("year", "WinterAvg.Shifted")], by = c("year"), all.x = TRUE)
colnames(data_aea)[colnames(data_aea)=="WinterAvg.Shifted"] <- "amo_winter" #change the colname
data_aea
```
ok and write a csv
```{r}
write.csv(data_aea,"../output/bio/data_aea_cell_amo.csv", row.names = FALSE)
```

### 2018-09-19 src/environmental_data_preperation.Rmd
amo data extraction

Now subset the dataset to start in 1997 (the capelin data starts in 1998, but you want some previous time period stuff)
```{r}
amoraw <- amoraw[amoraw$year >= 1997, ]
head(amoraw)
```

>>>>

Now subset the dataset to start in 1996 (the capelin data starts in 1998, but you want some previous time period stuff)
```{r}
amoraw <- amoraw[amoraw$year >= 1996, ]
head(amoraw)
```
